<!--
%\VignetteEngine{knitr}
%\VignetteIndexEntry{tximport}
-->

# tximport: import and summarize transcript-level estimates for gene-level analysis

## kallisto

We start with some kallisto TSV files containing transcript
abundance estimates. The pipeline will be nearly
identical for other quantitation tools, which are shown below.
First, read in some kallisto example files:

```{r}
library(tximportData)
dir <- system.file("extdata", package="tximportData")
list.files(dir)

samples <- read.table(file.path(dir,"samples.txt"), header=TRUE)
files <- file.path(dir,"kallisto", samples$run, "abundance.tsv")
names(files) <- paste0("sample",1:6)
```

Transcripts need to be associated with gene IDs for summarization.
If that information is present in the files, we can skip this step.
But for kallisto and Salmon, the files just provide the transcript ID.
So we first make a `data.frame` with two columns: gene ID (column 1)
and transcript ID (column 2).
The column names are not relevant but this column order must be used.
This can be accomplished from a *TxDb* object and the `select` function. 
(TODO: show example)

```{r}
gene2tx <- read.csv(file.path(dir, "gene2tx.csv"))
head(gene2tx)
```

The *tximport* package has a single function for importing transcript-level estimates.
The `type` argument is used to specify what software was used for estimation
("kallisto", "salmon", "rsem" implemented so far).
A simple list with matrices, "abundance", "counts", and "length", is returned.
The "length" matrix can be used to generate an offset matrix for downstream
gene-level differential analysis of count matrices.

While *tximport* works without any dependencies, it is much faster to
read in files using the *readr* package (version >= 0.2.2).
To use this, we pass the `read_tsv` function to `tximport`.

```{r}
library(tximport)
library(readr)
txi <- tximport(files, type="kallisto", gene2tx=gene2tx, reader=read_tsv)
names(txi)
head(txi$counts)
```

We could alternatively generate counts from abundances, 
using the argument `countsFromAbundance`,
scaled to library size (`"scaledTPM"`) or additionally scaled 
using the average transcript length,
averaged over samples and to library size (`"lengthScaledTPM"`). 
Using either of these approaches, the counts are not correlated 
with length, and so the length matrix does not need to be provided as an offset
for downstream analysis packages.

We can avoid gene-level summarization by setting `txOut=TRUE`.

## Salmon / Sailfish

Salmon or Sailfish `quant.sf` files can be imported by setting type to
`"salmon"` or `"sailfish"`.

```{r}
files <- file.path(dir,"salmon", samples$run, "quant.sf")
names(files) <- paste0("sample",1:6)
txi.salmon <- tximport(files, type="salmon", gene2tx=gene2tx, reader=read_tsv)
head(txi.salmon$counts)
```

## RSEM

```{r}
files <- file.path(dir,"rsem", samples$run, paste0(samples$run, ".genes.results"))
names(files) <- paste0("sample",1:6)
txi.rsem <- tximport(files, type="rsem", reader=read_tsv)
head(txi.rsem$counts)
```

## Import with edgeR, DESeq2, limma-voom

An example of creating a `DGEList` for use with edgeR:

```{r, results="hide", messages=FALSE}
library(edgeR)
```

```{r}
cts <- txi$counts
normMat <- txi$length
normMat <- normMat / exp(rowMeans(log(normMat)))
library(edgeR)
o <- log(calcNormFactors(cts/normMat)) + log(colSums(cts/normMat))
y <- DGEList(cts)
y$offset <- t(t(log(normMat)) + o)
# y is now ready for estimate dispersion functions
# see edgeR User's Guide
```

An example of creating a `DESeqDataSet` for use with DESeq2
(requires R-devel, Bioconductor 3.3 and DESeq2 version >= 1.11.6, or
you can source the function from the 
[development branch](https://github.com/Bioconductor-mirror/DESeq2/blob/master/R/AllClasses.R#L318-L333)).

```{r, results="hide", messages=FALSE}
library(DESeq2)
```

The user should make sure the rownames of `sampleTable` align with the
colnames of `txi$counts`, if there are colnames. The best practice is
to read `sampleTable` from a CSV file, and to construct `files` from a
column of `sampleTable` before calling *tximport*.

```{r}
sampleTable <- data.frame(condition=factor(rep(c("A","B"),each=3)))
rownames(sampleTable) <- colnames(txi$counts)
```

```{r}
dds <- DESeqDataSetFromTximport(txi, sampleTable, ~ condition)
# dds is now ready for DESeq()
# see DESeq2 vignette
```

An example for use with limma-voom:

```{r}
library(limma)
design <- model.matrix(~ condition, data=sampleTable)
v <- voom(y, design)
# v is now ready for lmFit()
# see limma User's Guide
```

## Session info

```{r}
sessionInfo()
```
